### 2021-06-01 paper/medical_parameters.tex
Start on updating the medical parameters.

This section discusses the medical parameters used in the model, their sources and how
we arrived at the distributions used in the model.\footnotemark

@@

\subsubsection{Length of Presymptomatic Stage / Incubation Period}

Estimates of the incubation period usually give a range from 2 to 12 days. A meta
analysis by \citet{McAloon2020} comes to the conclusion that ``The incubation period
distribution may be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$
parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55),
respectively.'' For simplicity we discretize this distribution into four bins.

\subsubsection{Begin of Infectiousness}

The period between infection and onset of infectiousness is called latent or latency
period. However, the latency period is rarely given in epidemiological reports on
Covid-19. Instead, scientists and agencies usually report the incubation period, the
period from infection to the onset of symptoms. A few studies used measurements of virus
shedding to estimate infectiousness during the course of the disease. When measurements
started before the onset of symptoms the development of the viral load before symptoms
gives us an indication of number of days between the onset of infectiousness and
symptoms.

The European Centre for Disease Prevention and Control estimates that people become
infectious between one and two days before the symptoms set in. This is similar to
\citet{He2020} who estimate this to take 2.3 days and is in line with \citet{Peak2020}.

Given these numbers and the length of the incubation period we can calculate the latency
period for symptomatic people. To our knowledge no estimates for the latency period of
asymptomatic cases of COVID-19 exist. We assume it to be the same for symptomatic and
asymptomatic cases.

Thus, we arrive at the following distribution for latency periods: 40\% have one day.
35\% have two days. 20\% have three days and 5\% have 5 days.

@@

We assume that the duration of infectiousness is the same for both symptomatic and
asymptomatic individuals as evidence suggests little differences in the transmission
rates of SARS-CoV-2 virus between symptomatic and asymptomatic patients
(\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic
individuals are similar (\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}).

@@

For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8
days of infectiousness afterwards.\footnotemark Thus, we arrive at 0 to 13 days as the
range for infectiousness among individuals who become symptomatic (see also figure 5).
This duration range is very much in line with the meta-analysis’ reported evidence for
asymptomatic individuals (see their figure 1). Thus, we arrive at 0 to 13 days as the
range for infectiousness among individuals who become symptomatic. This duration range
is very much in line with the meta-analysis' reported evidence for asymptomatic
individuals.

@@

The data on how many percent of symptomatic patients will require ICU is pretty thin. We
rely on data by the US CDC (\citet{Stokes2020}) and
\href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the
OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the
probabilities of requiring intensive care per age group.

>>>>

This section discusses the medical parameters used in the model, their sources and how we
arrived at the distributions used in the model.\footnotemark See
Figure~\ref{fig:course_of_disease} for a summary of our disease progression model.

@@

\subsubsection{Duration until Infectiousness and Symptoms}

% Latency Period
The first medical parameter we need is the length of the period between infection and
the start of infectiousness, the so called latent period.
We infer it from two other measures that are more common in the medical literature:
Firstly, the time between infection and the onset of symptoms, the incubation period.
Secondly, the time between the start of infectiousness and the onset of symptoms.
We assume that the latency period is the same for symptomatic and asymptomatic individuals.

% Asymptomatic Share
Once individuals become infectious a share of them goes on to develop symptoms while
others remain asymptomatic.
We rely on data by \cite{Davies2020} for the age-dependent probability to develop
symptoms. It varies from 25\% for children and young adults to nearly 70\% for the
elderly.

% Incubation Period
The incubation period is usually estimated to be two to twelve days.
A meta analysis by \citet{McAloon2020} comes to the conclusion that
``The incubation period distribution may be modeled with a lognormal distribution with
pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50
(95\% CI 0.46 to 0.55), respectively.''
For simplicity we discretize this distribution into four bins.

% Length of the presymptomatic stage
The \href{European Centre for Disease Prevention and Control}{https://bit.ly/3pi18Ax}
reports that people become infectious between one and two days before symptoms start.\footnotemark

\footnotetext{
This is similar to \citet{He2020} and in line with
\citet{Peak2020}.}

Taking these estimates together, we arrive at a latent period of one to five days.

@@

We assume that the duration of infectiousness is the same for both symptomatic and
asymptomatic individuals as evidence suggests little differences in the transmission
rates between symptomatic and asymptomatic patients
(\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic
individuals are similar (\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}).

@@

For symptomatic cases they arrive at 0-5 days before symptom onset (figure 2) and 3-8
days of infectiousness afterwards.\footnotemark Thus, we arrive at 0 to 13 days as the
range for infectiousness among individuals who become symptomatic (see also figure 5).
This duration range is very much in line with the meta-analysis’ reported evidence for
asymptomatic individuals (see their figure 1).

@@

The data on how many percent of symptomatic patients will require ICU is pretty thin. We
rely on data by the US CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the
OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the
probabilities of requiring intensive care per age group.
